ForOpenCL: Transformations Exploiting Array Syntax in Fortran for Accelerator Programming

### 2011-11-15 papers/IJCSE/shallow-water.tex
starting point for journal paper due Dec. 12

These simple code segments show how the shallow water model is implemented in
standard Fortran using the data-parallel programming model described above.  The
resulting code is simple, concise, and easy to understand.  However it does
\emph{not} necessarily perform well because of the temporary array variables,
especially those produced by the function {\tt region\_cpy}.  This is generally true
of algorithms that use Fortran shift functions as well, as some Fortran
compilers (e.g., gfortran) do not generate optimal code for shifts.  We note (as
shown below) that these temporary array copies are replaced by scalars in the
transformed Fortran code so there are no performance penalties for using
data-parallel statements as outlined.  However, there is an increased memory
cost due to the double buffering required by the kernel execution semantics.

>>>>

These simple code segments show how the shallow water model is implemented in
standard Fortran using the data-parallel programming model described above.  The
resulting code is simple, concise, and easy to understand.  However it does
\emph{not} necessarily perform well when compiled for a traditional sequential
system because of suboptimal use of temporary array variables,
especially those produced by the function {\tt region\_cpy}.  This is generally true
of algorithms that use Fortran shift functions as well, as some Fortran
compilers (e.g., gfortran) do not generate optimal code for shifts.  We note (as
shown below) that these temporary array copies are replaced by scalars in the
transformed Fortran code so there are no performance penalties for using
data-parallel statements as outlined.  However, there is an increased memory
cost due to the double buffering required by the kernel execution semantics.

### 2011-11-15 papers/IJCSE/source-to-source.tex
starting point for journal paper due Dec. 12

This section provides an brief overview of the ForOpenCL transformations that take
Fortran elemental and pure procedures as input and generate OpenCL code.  Recall
that elemental functions are transformed to inline OpenCL functions and that
subroutines with the {\tt PURE} and {\tt KERNEL} compiler directive attributes
are transformed to OpenCL kernels.

>>>>

This section provides an brief overview of the ForOpenCL
transformations that take Fortran elemental and pure procedures as
input and generate OpenCL code.  Elemental functions are
transformed to inline OpenCL functions and subroutines with the
{\tt PURE} and {\tt KERNEL} compiler directive attributes are
transformed to OpenCL kernels.
